SHARP - II. Mass structure in strong lenses is not necessarily substructure: A flux ratio anomaly from an edge-on disc

### 2016-02-23 papers/B1555_EdgeOn/Resubmission_v2/referee_response_2.txt

REFEREE: The revised version of the paper still contains the unsubstantiated speculation pointed out in my previous report, with no additional supporting evidence.  I am puzzled by the authors' statement in their response that an analysis of the effects of disks on lens statistics would be beyond the scope of the paper, since this is literally the topic of the paper.  The main problem now is that the paper's title is misleading, since nothing in this work provides any reason to believe that there is a bias in observational measurements of dark matter substructure with gravitational lens flux ratios.

RESPONSE: The referee is correct to point out that our title may be misleading in implying that we intend to analyze the effects of disks on lens statistics. Thus, we change our title in order to focus on the disc effect itself. Also, we add more text in paper to show how we can be informed by a new model of a single system. That being said, discs with high inclination in lens system can significantly affect the lensing cross-section, as has been shown in Maller et al. 1997, Möller et al. 2003, and Keeton & Kochanek 1998.  Although the fraction of high-inclination spiral lenses may not be high in the general population, the current sample of lensed quasars that have flux-ratio anomalies is fairly small size (around 10 in all).  Therefore, if even a few systems within the sample have edge-on discs, these are sufficient to bias the substructure analysis. As previous work on flux ratios in lensed quasar systems did its analysis with standard SIE mass profiles, it is fair to say that the disc effect is a bias that been ignored for a long time. Moreover, from our high-resolution IR imaging of anomalous flux-ratio lenses, we know that B1555 is not the only system that has an edge-on disc across lensed images. Since this is the first time that an edge-on disc has been detected that can account for at least some part of observed flux ratio anomalies, we think it is reasonable to talk about this ‘potential bias’ in the discussion.

### 2016-02-23 papers/B1555_EdgeOn/B1555_letter.tex
Further update of the manuscript text

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  \textbf{Although a constraint on the disc mass from photometry is currently not avaliable due to the low signal-to-noise ratio in AO imaging and the lack of redshift information, our model indicates that the flux-ratio anomalies in B1555+375, at least, have a strong link to the presence of its edge-on disc. } Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomalies are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc. \textbf{Some of concerns may arise that an overestimated disc component leads to a model can explain the entire anomaly. One is saying can a spiral lens with massive disc not distort the arc close to the disc. J0820+4847, one of the SWELLS samples, is an edge-on lens has an arc right across its disc. The SWELLS studies indicate it has 15 per cent of disc mass within the Einstein radius. Other SWELLS lenses have disc mass fraction range from 5 per cent to 30 per cent \citep{swells1,swells3,swells6}. These numbers tell that the disc in our best-fit model of B1555+375 still fits into the nature of general spiral lenses.}  However, it may be possible to independently constrain the mass
of the disc through kinematic measurements (most likely from
mm-wavelength resolved emission line observations), which would test
whether the disc actually makes a sufficient contribution to the mass
model \citep[e.g. as in the case of B1933+503;][]{suyu12}. Also, in
the blue contours of Fig.~\ref{fig:3panel}a, B1555+375 shows evidence
for an extended $\sim$ 100 mas gravitational arc at radio wavelengths
that crosses the edge-on disc. Further high resolution imaging with
very long baseline interferometry will provide the observational
constraints required to fully test whether the disc is sufficient to
cause the anomaly.

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. \textbf{\citet{Maller97,Kee98, Moller03} have shown that a disc with high inclination angle can significantly affects the lensing cross-section. While the current sample of lensed quasars with flux ratio anomalies is fairly small, few systems with edge-on disc may be sufficient to bring a bias into the substructure analysis. Moreover, near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component detection close to merging images among the current sample of flux ratio anomalous systems; these gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have similar
links between flux-ratio anomalies and disc components. This potential bias brought by edge-on disc may be avoided by post-selection as the sample size grows in the future. Therefore, it is critical to obtain high-resolution imaging like SHARP of target lenses to detect `sharp' mass features such as edge-on disks or even massive luminous satellites.} 
% A number of lens systems with significant disc components arealready known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}. 
 This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.

>>>>

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  \textbf{Although a constraint on the disc mass from photometry is currently not avaliable due to the low signal-to-noise ratio in the AO imaging and the lack of redshift information, our model indicates that the flux-ratio anomalies in B1555+375, at least, have a strong link to the presence of its edge-on disc. } Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomalies are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc. 
\textbf{That is not to imply this is a unique model, but rather that the
disc component is a possible explanation of the anomaly.  We note that
one of the SWELLS systems, J0820+4847, has an edge-on disc that directly
crosses but does not significantly distort the arc of lensed emission in the system \citep{swells1}. The disc in this system
has been shown to comprise 15 per cent of the mass within the Einstein
radius, comparable to our results on B1555+375 and also fully consistent with
the 5 to 30 per cent contributions of discs in the other lenses in the 
SWELLS sample \citep{swells1,swells3,swells6}.}

In future it may be possible to independently constrain the mass
of the disc through kinematic measurements (most likely from
mm-wavelength resolved emission line observations), which would test
whether the disc actually makes a sufficient contribution to the mass
model \citep[e.g. as in the case of B1933+503;][]{suyu12}. Also, in
the blue contours of Fig.~\ref{fig:3panel}a, B1555+375 shows evidence
for an extended $\sim$ 100 mas gravitational arc at radio wavelengths
that crosses the edge-on disc. Further high resolution imaging with
very long baseline interferometry will provide the observational
constraints required to fully test whether the disc is sufficient to
cause the anomaly.

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. \textbf{Special care may be needed for small-separation lens systems, in which the lensing galaxy is likely to be a spiral [CITE Turner, Ostriker, Gott].  This is because \citet{Maller97}, \citet{Kee98}, and \citet{Moller03} have shown that a disc with high inclination angle can significantly affect the lensing cross-section for such systems. The current sample of lensed quasars with radio-wavelength flux ratio measurements is small.  Therefore, having even a few systems with edge-on discs may be sufficient to introduce a bias into the substructure analysis. In fact, our near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component that crosses the merging lensed images.  These gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have similar
links between flux-ratio anomalies and disc components. This potential bias brought by edge-on disc may be avoided by post-selection as the sample size grows in the future. Therefore, it is critical to obtain high-resolution imaging like SHARP of target lenses to detect `sharp' mass features such as edge-on disks or even massive luminous satellites.} 
% A number of lens systems with significant disc components arealready known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}. 
 This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, Euclid, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.
